---
title: "Regularisation tests"
author: "Cameron Roach"
date: "03/03/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(lightgbm)
library(podEnergyComp)

demand.data <- load_demand_data() %>% 
  filter(!is.na(demand_mw))
pv.data <- load_pv_data() %>% 
  filter(!is.na(pv_power_mw))

cv_iter <- 5
# FIXME: calculate different start_date for PV OR change this to date instead of datetime. May need to work on cv_ts_folds()
start_date <- max(demand.data$datetime) + minutes(30) - days(7*cv_iter)
# start_date <- ymd("2018-10-15") - days(7*cv_iter)
```

# Demand

## Number of leaves and learning rate

### All periods modelled together

```{r}
demand.cv_df <- cross_df(
  list(
    "lr" = c(0.05, 0.1, 0.15),
    "nl" = c(8, 10, 12, 14),
    "cv" = cv_ts_folds(
      demand.data$datetime,
      start_date = start_date,
      horizon = 7,
      iterations = cv_iter
    ),
    "pred" = NA
  )
)

for (i in 1:nrow(demand.cv_df)) {
  print(paste0(i, "/", nrow(demand.cv_df), " : ", Sys.time()))
  pred <- pred_demand(
    select(demand.data, -datetime),
    demand.cv_df$cv[[i]]$train,
    demand.cv_df$cv[[i]]$test,
    nrounds = 500,
    num_leaves = demand.cv_df$nl[i],
    learning_rate = demand.cv_df$lr[i],
    obj = "regression",
    metric = "regression",
  )
  
  demand.cv_df$pred[i] <- 
    list(
      tibble(
        datetime = getElement(demand.data[demand.cv_df$cv[[i]]$test,],
                              "datetime"),
        demand_mw = getElement(demand.data[demand.cv_df$cv[[i]]$test,],
                               "demand_mw"),
        pred_demand_mw = pred
      ) %>% 
        mutate(idx = row_number())  # useful for plotting
    )
}

demand.results <- demand.cv_df %>% 
  filter(!is.na(pred)) %>% 
  unnest(pred) %>% 
  mutate(e = demand_mw - pred_demand_mw,
         ae = abs(e)) %>% 
  group_by(lr, nl) %>% 
  summarise(mae = mean(ae),
            rmse = sqrt(mean(e^2)),
            .groups = "drop") %>% 
  arrange(rmse)
demand.results

demand.results %>% 
  mutate(lr = factor(lr),
         nl = factor(nl)) %>% 
  ggplot(aes(x = lr, y = nl, fill = rmse)) +
  geom_tile() +
  geom_text(aes(label = round(rmse, 4))) +
  scale_fill_viridis_c(direction = -1) +
  labs(title = "RMSE values",
       subtitle = paste0("Start date: ", start_date, ", iter: ", cv_iter))

best_param <- demand.results %>% 
  filter(rmse == min(rmse)) %>% 
  select(lr, nl)

demand.cv_df %>% 
  filter(lr == best_param$lr,
         nl == best_param$nl) %>% 
  mutate(iter = row_number()) %>% 
  unnest(pred) %>% 
  select(iter, idx, ends_with("mw")) %>% 
  pivot_longer(cols = ends_with("_mw")) %>% 
  ggplot(aes(x = idx, y = value, colour = name)) + 
  geom_line() + 
  facet_wrap(~iter) +
  labs(title = "Predictions and actuals for cross validation iterations",
       subtitle = paste0("lr = ", best_param$lr, ", nl = ", best_param$nl))

demand.cv_df %>% 
  filter(lr == best_param$lr,
         nl == best_param$nl) %>% 
  mutate(iter = row_number()) %>% 
  unnest(pred) %>% 
  ggplot(aes(x = demand_mw, y = pred_demand_mw)) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, colour = "red", alpha = 0.5) +
  facet_wrap(~iter) +
  labs(title = "Predictions and actuals for cross validation iterations",
       subtitle = paste0("lr = ", best_param$lr, ", nl = ", best_param$nl))
```

Best result is `lr = 0.1` and `nl = 8`.


### Periods modelled separately


```{r eval = F}
demand.cv_df <- cross_df(
  list(
    "lr" = c(0.03, 0.05),
    "nl" = c(6, 8),
    "cv" = cv_ts_folds(
      demand.data$datetime,
      start_date = start_date,
      horizon = 7,
      iterations = cv_iter
    ),
    "pred" = NA
  )
)

for (i in 1:nrow(demand.cv_df)) {
  print(paste0(i, "/", nrow(demand.cv_df), " : ", Sys.time()))
  pred <- pred_demand_period(
    select(demand.data, -datetime),
    demand.cv_df$cv[[i]]$train,
    demand.cv_df$cv[[i]]$test,
    nrounds = 200,
    min_data_in_leaf = 2,
    num_leaves = demand.cv_df$nl[i],
    learning_rate = demand.cv_df$lr[i],
    obj = "regression",
    metric = "regression",
  )
  
  demand.cv_df$pred[i] <- 
    list(
      tibble(
        datetime = getElement(demand.data[demand.cv_df$cv[[i]]$test,],
                              "datetime"),
        demand_mw = getElement(demand.data[demand.cv_df$cv[[i]]$test,],
                               "demand_mw"),
        pred_demand_mw = pred
      ) %>% 
        mutate(idx = row_number())  # useful for plotting
    )
}

demand.results <- demand.cv_df %>% 
  filter(!is.na(pred)) %>% 
  unnest(pred) %>% 
  mutate(e = demand_mw - pred_demand_mw,
         ae = abs(e)) %>% 
  group_by(lr, nl) %>% 
  summarise(mae = mean(ae),
            rmse = sqrt(mean(e^2)),
            .groups = "drop") %>% 
  arrange(rmse)
demand.results

demand.results %>% 
  mutate(lr = factor(lr),
         nl = factor(nl)) %>% 
  ggplot(aes(x = lr, y = nl, fill = rmse)) +
  geom_tile() +
  geom_text(aes(label = round(rmse, 4))) +
  scale_fill_viridis_c(direction = -1) +
  labs(title = "RMSE values",
       subtitle = paste0("Start date: ", start_date, ", iter: ", cv_iter))

best_param <- demand.results %>% 
  filter(rmse == min(rmse)) %>% 
  select(lr, nl)

demand.cv_df %>% 
  filter(lr == best_param$lr,
         nl == best_param$nl) %>% 
  mutate(iter = row_number()) %>% 
  unnest(pred) %>% 
  select(iter, idx, ends_with("mw")) %>% 
  pivot_longer(cols = ends_with("_mw")) %>% 
  ggplot(aes(x = idx, y = value, colour = name)) + 
  geom_line() + 
  facet_wrap(~iter) +
  labs(title = "Predictions and actuals for cross validation iterations",
       subtitle = paste0("lr = ", best_param$lr, ", nl = ", best_param$nl))

demand.cv_df %>% 
  filter(lr == best_param$lr,
         nl == best_param$nl) %>% 
  mutate(iter = row_number()) %>% 
  unnest(pred) %>% 
  ggplot(aes(x = demand_mw, y = pred_demand_mw)) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, colour = "red", alpha = 0.5) +
  facet_wrap(~iter) +
  labs(title = "Predictions and actuals for cross validation iterations",
       subtitle = paste0("lr = ", best_param$lr, ", nl = ", best_param$nl))
```




## Compare to lambda_l1

The above numbers are consistently better (and faster to run) than regularization with only `lambda_l1`. Can compare by running the below code.

```{r eval=FALSE}
# cv2_df <- 
#   cross_df(
#     list(
#       "nl" = 8L,
#       "lr" = 0.1,
#       "lambda_l1" = c(0.1, 1.0, 10),
#       "cv" = cv_ts_folds(
#         demand.data$datetime,
#         start_date = start_date,
#         horizon = 7,
#         iterations = cv_iter
#       ),
#       "pred" = NA
#     )
#   ) %>%
#   bind_rows(
cv2_df <- cross_df(
  list(
    nl = 31L,
    lr = 0.03,
    lambda_l1 = exp(seq(-1,4)),
    "cv" = cv_ts_folds(
      demand.data$datetime,
      start_date = start_date,
      horizon = 7,
      iterations = cv_iter
    ),
    "pred" = NA
  )
)

for (i in 1:nrow(cv2_df)) {
  print(paste0(i, "/", nrow(cv2_df), " : ", Sys.time()))
  pred <- pred_demand(
    demand.data[,-1],
    cv2_df$cv[[i]]$train,
    cv2_df$cv[[i]]$test,
    nrounds = 250,
    num_leaves = cv2_df$nl[i],
    learning_rate = cv2_df$lr[i],
    lambda_l1 = cv2_df$lambda_l1[i],
    obj = "regression_l1",
    metric = "regression_l1",
  )
  
  cv2_df$pred[i] <- 
    list(
      tibble(
        datetime = getElement(demand.data[cv2_df$cv[[i]]$test,], "datetime"),
        demand_mw = getElement(demand.data[cv2_df$cv[[i]]$test,], "demand_mw"),
        pred_demand_mw = pred
      ) %>% 
        mutate(idx = row_number())  # useful for plotting
    )
}

results2 <- cv2_df %>% 
  filter(!is.na(pred)) %>% 
  unnest(pred) %>% 
  mutate(e = demand_mw - pred_demand_mw,
         ae = abs(e)) %>% 
  group_by(lambda_l1, nl, lr) %>% 
  summarise(mae = mean(ae),
            rmse = sqrt(mean(e^2)),
            .groups = "drop") %>% 
  arrange(rmse)
results2
```


# PV

PV neds to be explored more. Optimal nl and lr paramaters give pv power predictions in the middle of the night. Sticking with L1 optimisation for the moment.

```{r eval=FALSE}
pv.cv_df <- cross_df(
  list(
    "lr" = c(0.05, 0.1, 0.2, 0.3),
    "nl" = c(2, 4, 8, 16, 24),
    "cv" = cv_ts_folds(
      pv.data$datetime,
      start_date = start_date,
      horizon = 7,
      iterations = cv_iter
    ),
    "pred" = NA
  )
)

for (i in 1:nrow(pv.cv_df)) {
  print(paste0(i, "/", nrow(pv.cv_df), " : ", Sys.time()))
  pred <- pred_pv(
    select(pv.data, -datetime),
    pv.cv_df$cv[[i]]$train,
    pv.cv_df$cv[[i]]$test,
    nrounds = 500,
    num_leaves = pv.cv_df$nl[i],
    learning_rate = pv.cv_df$lr[i],
    obj = "regression",
    metric = "regression",
  )
  
  pv.cv_df$pred[i] <- 
    list(
      tibble(
        datetime = getElement(pv.data[pv.cv_df$cv[[i]]$test,], "datetime"),
        pv_power_mw = getElement(pv.data[pv.cv_df$cv[[i]]$test,], "pv_power_mw"),
        pred_pv_power_mw = pred
      ) %>% 
        mutate(idx = row_number())  # useful for plotting
    )
}
```


### Results

```{r eval=FALSE}
pv.results <- pv.cv_df %>% 
  filter(!is.na(pred)) %>% 
  unnest(pred) %>% 
  mutate(e = pv_power_mw - pred_pv_power_mw,
         ae = abs(e)) %>% 
  group_by(lr, nl) %>% 
  summarise(mae = mean(ae),
            rmse = sqrt(mean(e^2)),
            .groups = "drop") %>% 
  arrange(rmse)
pv.results

pv.results %>% 
  mutate(lr = factor(lr),
         nl = factor(nl)) %>% 
  ggplot(aes(x = lr, y = nl, fill = rmse)) +
  geom_tile() +
  scale_fill_viridis_c(direction = -1)

best_param <- pv.results %>% 
  filter(rmse == min(rmse)) %>% 
  select(lr, nl)

pv.cv_df %>% 
  filter(lr == best_param$lr,
         nl == best_param$nl) %>% 
  mutate(iter = row_number()) %>% 
  unnest(pred) %>% 
  select(iter, idx, ends_with("mw")) %>% 
  pivot_longer(cols = ends_with("_mw")) %>% 
  ggplot(aes(x = idx, y = value, colour = name)) + 
  geom_line() + 
  facet_wrap(~iter) +
  labs(title = "Predictions and actuals for cross validation iterations",
       subtitle = paste0("lr = ", best_param$lr, ", nl = ", best_param$nl))

pv.cv_df %>% 
  filter(lr == best_param$lr,
         nl == best_param$nl) %>% 
  mutate(iter = row_number()) %>% 
  unnest(pred) %>% 
  ggplot(aes(x = pv_power_mw, y = pred_pv_power_mw)) + 
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, colour = "red", alpha = 0.5) +
  facet_wrap(~iter) +
  labs(title = "Predictions and actuals for cross validation iterations",
       subtitle = paste0("lr = ", best_param$lr, ", nl = ", best_param$nl))
```

Best result is `lr = 0.1` and `nl = 4`.
